---
title: "Rent and Inventory SARIMA and ETS"
output: html_notebook
---

```{r}
rm(list=ls())
library(tidyverse)
library(tseries)
library(forecast)
library(TSA)
library(dplyr)
```

### Load and inspect data

```{r}
# source: https://streeteasy.com/blog/data-dashboard/?agg=Total&metric=Inventory&type=Rentals&bedrooms=Any%20Bedrooms&property=Any%20Property%20Type&minDate=2010-01-01&maxDate=2022-12-01&area=Flatiron,Brooklyn%20Heights

# note that we can also get breakdown by bedrooms if we want

# set up generalizable path without local path hard coded
base_path <- system('git rev-parse --show-toplevel', intern = T)

# rental index - one for each of four real boroughs
# note: we won't use the rental index as it is already smoothed via ARIMA
rental_index_raw <- read.csv(paste0(base_path, '/data/rentalIndex_All.csv'))
# median asking rent - four real boroughs + 155 neighborhoods
median_rent_raw <- read.csv(paste0(base_path, '/data/medianAskingRent_All.csv'))
# rental inventory - four real boroughs + 155 neighborhoods
inventory_raw <- read.csv(paste0(base_path, '/data/rentalInventory_All.csv'))

```

```{r}
rental_index_manhattan <- ts(rental_index_raw$Manhattan,
                             start=c(2007,1),
                             frequency=12)

tsdisplay(rental_index_manhattan)

median_rent_manhattan <- median_rent_raw %>% 
  filter(areaName=="Manhattan") %>%
  unlist() %>% as.numeric() %>% 
  na.omit() %>% 
  ts(start=c(2010,1),frequency=12)

tsdisplay(median_rent_manhattan)

inventory_manhattan <- inventory_raw %>% 
  filter(areaName=="Manhattan") %>% 
  unlist() %>% as.numeric() %>% 
  na.omit() %>%
  ts(start=c(2010,1),frequency=12)

tsdisplay(inventory_manhattan)

rent_train <- window(median_rent_manhattan, start=2010, end=c(2020,1))
inventory_train <- window(inventory_manhattan, start=2010, end=c(2020,1))
rent_test <- window(median_rent_manhattan, start=c(2020,2), end=c(2022,12))
inventory_test <- window(inventory_manhattan, start=c(2020,2), end=c(2022,12))

# Box-Cox transformed
tsdisplay(BoxCox(rent_train,lambda='auto'))
tsdisplay(BoxCox(inventory_train,lambda='auto'))
```

```{r}
#save(inventory_manhattan,file='data/inventory_manhattan.RData')
#save(median_rent_manhattan,file='data/median_rent_manhattan.RData')
```


```{r}
plot((rent_train-mean(rent_train))/var(rent_train)/30)
lines((inventory_train-mean(inventory_train))/var(inventory_train),col='red')
```

```{r}
rent_stationary <- diff(BoxCox(rent_train,lambda=0),differences=1,lag=1)
inventory_stationary <- diff(BoxCox(inventory_train,lambda=0),differences=1,lag=1)

plot((rent_stationary-mean(rent_stationary))/var(rent_stationary))
lines(((inventory_stationary-mean(inventory_stationary)))/var(inventory_stationary),col='red')
```


### SARIMA

```{r}
# median rent SARIMA
rent_lambda = BoxCox.lambda(rent_train)
rent_sarima <- auto.arima(rent_train,lambda=rent_lambda,trace=TRUE)
checkresiduals(rent_sarima)
summary(rent_sarima)
```
Best rent SARIMA: nonseasonal first difference, seasonal AR(2), no drift. This is a parsimonous model with white noise residuals. Adding drift, nonseasonal AR(1) and seasonal MA(1) only slightly increase AICc. RMSE 31.8, MAE 24,08806, MAPE 0.7533599.


```{r}
inventory_lambda = 0  # Note: As determined in Wesley's code, we will just use a log transform for the inventory series.
inventory_sarima <- auto.arima(inventory_train,lambda=inventory_lambda,trace=TRUE)
checkresiduals(inventory_sarima)
summary(inventory_sarima)
```

Best inventory SARIMA: nonseasonal first difference, seasonal first-differenced MA(1). Residuals are mostly white noise - possibly some quarterly / two-year autocorrelations, but might also just be multiple testing (Ljung-Box doesn't reject). Similarly to above, there are a number of models with very close AICc, though none with drift. RMSE 673.5361, MAE 481.7169, MAPE 2.686797.


### ETS

```{r}

rent_ets <- ets(rent_train, lambda = rent_lambda)  # note that we get AAA without lambda
summary(rent_ets)
checkresiduals(rent_ets)

inventory_ets <- ets(inventory_train)  # don't force lambda, because that forces additive model
summary(inventory_ets)
checkresiduals(inventory_ets)
```
Rent ETS: AAA (additive error, trend, seasonality).
    alpha = 0.9996 
    beta  = 2e-04 
    gamma = 3e-04 
    phi   = 0.9744 
Dominated by error, with very little trend or seasonality. Only slightly damped. Residuals mostly look like white noise, maybe 1.5 year seasonality, but Ljung-Box rejects at high significance level. Training RMSE 29.52856, MAE 22.48471, MAPE 0.7053597.

Inventory ETS: MAM (multiplicative error, additive trend, multiplicative seasonality).
    alpha = 0.9884 
    beta  = 0.0875 
    gamma = 1e-04 
    phi   = 0.98
Dominated by error, with some trend contribution and very little seasonality. Only slightly damped. Residuals mostly look like white noise but Ljung-Box rejects at high significance level. RMSE 621.3574, MAE 469.1133, MAPE 2.705515.


### Cross-validation

```{r}
source('cross_validation.R')
```

```{r}
# rent
output_arima <- cross.validate(rent_SARIMA_model,rent_train,80,12)
output_ets <- cross.validate(rent_ets_model,rent_train,80,12)
generate.plots(output_arima,output_ets)
```

```{r}
# inventory
output_arima <- cross.validate(inventory_SARIMA_model,inventory_train,80,12)
output_ets <- cross.validate(inventory_ets_model,inventory_train,80,12)
generate.plots(output_arima,output_ets)
```

### Intervention analysis

```{r}
# rent ARIMA
rent_covid <- ts(median_rent_manhattan[(length(rent_train)+1):length(median_rent_manhattan)],frequency=12,start=c(2020,2))
plot(forecast(rent_sarima,length(rent_covid)))
lines(rent_covid)
rent_effect <- rent_covid-forecast(rent_sarima,length(rent_covid))$mean
plot(rent_effect,main='Estimated COVID Effect on Rent (via ARIMA)')
mean(rent_effect)
abline(h=0)
```

```{r}
# inventory ARIMA
inventory_covid <- ts(inventory_manhattan[(length(rent_train)+1):length(inventory_manhattan)],frequency=12,start=c(2020,2))
plot(forecast(inventory_sarima,length(inventory_covid)))
lines(inventory_covid)
inventory_effect <- inventory_covid-forecast(inventory_sarima,length(inventory_covid))$mean
plot(inventory_effect,main='Estimated COVID Effect on Inventory (via ARIMA)')
mean(inventory_effect)
abline(h=0)
```

```{r}
# rent ETS
plot(forecast(rent_ets,length(rent_covid)))
lines(rent_covid)
rent_effect_ets <- rent_covid-forecast(rent_ets,length(rent_covid))$mean
plot(rent_effect_ets,main='Estimated COVID Effect on Rent (via ETS)')
mean(rent_effect_ets)
abline(h=0)
```


```{r}
# inventory ETS
plot(forecast(inventory_ets,length(inventory_covid)))
lines(inventory_covid)
inventory_effect_ets <- inventory_covid-forecast(inventory_ets,length(inventory_covid))$mean
plot(inventory_effect_ets,main='Estimated COVID Effect on Inventory (via ETS)')
mean(inventory_effect_ets)
abline(h=0)
```


```{r}
intervention_vec <- rep(c(0,1), c(122, 34))
# allow free estimation
summary(auto.arima(median_rent_manhattan,xreg=intervention_vec,lambda=rent_lambda))
summary(auto.arima(inventory_manhattan,xreg=intervention_vec,lambda=inventory_lambda))
# specify models as above
summary(Arima(median_rent_manhattan,c(0,1,0),c(2,0,0),xreg=intervention_vec,lambda=rent_lambda))
summary(Arima(inventory_manhattan,order=c(0,1,0),seasonal=c(0,1,1),xreg=intervention_vec,lambda=inventory_lambda))
```

























